---
title: "assignment_3"
output: html_document
---



## Question 1
```{r}
library(RCurl)
library(dplyr)

data <- getURL('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')
df <- read.csv(text = data)[2]
filter(df, grepl('DATA|STATISTICS', Major))
df_2 <- unlist(strsplit(data, ","))
grep(pattern = 'data|statistics', df_2, value = TRUE, ignore.case = TRUE)
```

## Question 2
```{r}
library(stringr)
data_2 <- readLines("data.txt")
data_2 <- unlist(str_extract_all(data_2, "[[:alpha:] ]{2,}"))
data_2 <- trimws(data_2, which = c("both", "left", "right"), whitespace = "[ \t\r\n]")
stri_remove_empty(data_2, na_empty = FALSE)
```
## Question 3

a = "(.)\1\1"  

This looks for strings where three letters reapeat themselves consecutively in PCRE but will not work in POSIX without another backslash.  

b = "(.)(.)\\2\\1"  

This code pulls out groups of 4 letter palendromes (soos = TRUE, 1221 = TRUE) in the POSIX editor. 

c = "(..)\1"  

Find groups of two letter that repeat themselves consecutively in PCRE but not work in POSIX withour another backslash.  

d = "(.).\\1.\\1"  

This extract a group of 5 chacters where at least the first, second, third, and fifth letter are all the same in POSIX. (ooomo = TRUE)

e = "(.)(.)(.).*\\3\\2\\1"

The core code finds a 7 level palendrome similar to the problem in (b) The middle chacter can be anything and not limited to letters. In this case the * designates there can be any number of chacters that occor in between. (racecar = TRUE, race hsdk234 car = TRUE) 


## Question 4

```{r}
test <- c('g33ogh', 'mohomsdjfj', 'doodomjj moom  doodomjj moo', 'jddj', 'abc     cba', 'dddwddd', 'racecar')
str_extract(test, '^\\w') == str_extract(test, '\\w$')
str_extract_all(test, '(\\w).*\\1.*\\1')
str_extract_all('chruch','(..).*\\1')
```





